Lifestyle Choice Matters: Exploring Population Health Disparities and the Effect of Tobacco and Alcohol on Life Expectancy

### 2023-11-06 outputs/paper/paper.qmd
Correct spelling errors

---
title: "Lifestyle Choice Matters: Exploring Population Health Disparities and the Effect of Tobacco and Alcohol on Life Expectancy"
author: "Yufei Liu"
thanks: "Code and data are available at: https://github.com/Florence-Liu/life_expectancy"
date: today
date-format: long
abstract: "Life expectancy is an important measurement of population health and there are many significant influencing factors such as genetic and lifestyle. This paper examines the relationship between the life expectancy, the prevalence of tobacco products, and the alcohol consumption using data from Global Health Observatory data repository from the WHO. Using regional comparisons, we find as prevalence of tobacco increases, the life expectancy at age 60 decreases while as alcohol consumption increases, the life expectancy increases. We also find regional, temporal, and sexual differences in life expectancy at age 60. Further work could consider more influencing varibles and look at more complex models to capture non-linear relationships."
format: pdf
number-sections: true
toc: false
bibliography: references.bib
---

@@

To justify our choice of including interaction terms based on observation, we will fit two models, model 1 contains interaction terms and model 2 does not include interaction terms. The full model is shown as

@@

Other than tobacco products, the effect of alcohol on health reveals a complex relationship. The full model we choose suggested that the alcohol consumption is associated with increased life expectancy at age 60. However, in the reduced model, when we do not consider sex impact on alcohol consumption, it shows a negative association. Based on our justification of taking into account sex interaction with alcohol consumption, it indicates an overall positive association. Additionally, evidence-based study has found that modest drinking has been approved to be protective against some diseases, but in most cases, alcohol could be seen as a health risk factor [@citeDrinking]. In our model, we did not specify alcohol consumption level, so it could be further investigated by creating a new dummy variable indicating different drinking levels.

>>>>

---
title: "Lifestyle Choice Matters: Exploring Population Health Disparities and the Effect of Tobacco and Alcohol on Life Expectancy"
author: "Yufei Liu"
thanks: "Code and data are available at: https://github.com/Florence-Liu/life_expectancy"
date: today
date-format: long
abstract: "Life expectancy is an important measurement of population health and there are many significant influencing factors such as genetic and lifestyle. This paper examines the relationship between the life expectancy, the prevalence of tobacco products, and the alcohol consumption using data from Global Health Observatory data repository from the WHO. Using regional comparisons, we find as prevalence of tobacco increases, the life expectancy at age 60 decreases while as alcohol consumption increases, the life expectancy increases. We also find regional, temporal, and sexual differences in life expectancy at age 60. Further work could consider additional influencing variables and look at more complex models to capture non-linear relationships."
format: pdf
number-sections: true
toc: false
bibliography: references.bib
---

@@

To justify our choice of including interaction terms based on observation, we will fit two models. Model 1 is the full model including interaction terms and model 2 is the reduced model without interaction terms. The full model is shown as

@@

Other than tobacco products, the effect of alcohol on health reveals a complex relationship. The full model we choose suggested that the alcohol consumption is associated with increased life expectancy at age 60. However, in the reduced model, when we do not consider sex impact on alcohol consumption, it shows a negative association. This is consistent with the data visualization that shows a Simpson's Paradox. Based on our justification of taking into account sex interaction with alcohol consumption, it indicates an overall positive association. Additionally, evidence-based study has found that modest drinking has been approved to be protective against some diseases, but in most cases, alcohol could be seen as a health risk factor [@citeDrinking]. In our model, we did not specify alcohol consumption level, so it could be further investigated by creating a new dummy variable indicating different drinking levels.
